---
title: "Design Homework 1"
author: "Rachel Jordan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Questions from Design and Analysis of Experiments, 8th edition, Douglas C. Montgomery.*

## Question 2.15

a)  Yes, the null hypothesis can be rejected because the CI does not include 0 and the p-value is below 0.05.

b)  This is a two-sided test.

c)  Yes, the null hypothesis could be rejected because the CI does not include 2.

d)  Yes, because both bounds of the CI are below 2. You can answer this question without doing any additional calculations by looking at the CI.

e)  

```{r}
tvalue <- 1.697
difference <- -2.33341
pooledstdev <- 2.1277
root <- sqrt((1/20)+(1/20))

difference + tvalue*pooledstdev*root
```

f\.

```{r}
test_stat_f <- (difference - 2)/(pooledstdev*root)

pt(test_stat_f,38,lower.tail=T)
```

The p-value is 7.094e-08.

    7.09418e-08

## Question 2.28

a\.

```{r}
var1 <- 101.17
var2 <- 94.73
fvalue <- var1/var2
ftablevalue <- 3.5005
fvalue > ftablevalue
```

The calculated F-value is smaller than the value from the F table, so we cannot reject the null hypothesis. The variances are not significantly different.

b\.

```{r}
#pooled t-test
y1bar <- 12.5
y2bar <- 10.2
n1 <- 8
n2 <- 9
spooled <- sqrt(((n1-1)*var1 + (n2-1)*var2) / (n1+n2-2))
teststat228 <- (y1bar-y2bar)/(spooled*sqrt((1/n1)+(1/n2)))
abs(teststat228) > qt(p=0.05,df=(n1+n2-2),lower.tail=F)
```

The test statistic is less than the critical value, so we cannot reject the null. We cannot say that the two sample means are significantly different. Of course this is assuming that these samples are normally distributed.

## Question 2.31

```{r}
wafers <- as.numeric(c(5.34,6.00,5.97,5.25,6.65,7.55,7.35,6.35,4.76,5.54,5.44,4.61,5.98,5.62,4.39,6.00,7.25,6.21,4.98,5.32))
```

a)  

```{r}
require(stats)
wafer_var <- var(wafers)
n_231 <- length(wafers)
lower_denom <- 32.85
upper_denom <- 8.91
lower_bound <- ((n_231-1)*wafer_var)/lower_denom
upper_bound <- ((n_231-1)*wafer_var)/upper_denom
cat(lower_bound,upper_bound)
```

A 95% CI for the variance is (0.457,1.686).

b)  

```{r}
teststat_231 <- ((n_231-1)*wafer_var)/1
teststat_231 > lower_denom
teststat_231 < upper_denom
```

We cannot reject the null hypothesis at alpha=0.05. The true value of the variance could be 1.

c)  We are assuming that the data are normally distributed for this test because the chi-square table is derived from the assumption of normality of the parent distribution (Davies, 1956 -- cited in our textbook). Square of normal is a (scaled if not 0,1) chi-square, so you need a normal distribution to get your pivot for these inferences.

d)  

```{r}
qqnorm(wafers)
```

The data don't look heinously abnormal. This can be confirmed with a Shapiro-Wilk test that has an insignificant p-value, indicating normality. We can treat the data as normal.

```{r}
shapiro.test(wafers)
```

## Question 2.36

```{r}
form_1 <- c(206,188,205,187,193,207,185,189,192,210,194,178)
form_2 <- c(177,197,206,201,176,185,200,197,198,188,189,203)

t.test(form_1,form_2,var.equal=T,mu=3,alternative="greater")
```

The p-value is insignificant and the confidence interval includes 3 and some values below 3, which means that we do not have sufficient evidence to conclude that the mean deflection temperature under load for formulation 1 exceeds that of formulation 2 by at least 3F.
