---
title: "Jordan_Hw5"
author: "Rachel Jordan"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 6.8

```{r}
#input data
A_68 <- c(-1, 1, -1,1)
B_68 <- c(-1, -1, 1, 1)
I_68 <- c(21,37,25,31)
II_68 <- c(23,38,24,29)
III_68 <- c(20,35,29,30)
IV_68 <- c(22,39,26,34)
V_68 <- c(28,38,25,33)
VI_68 <- c(26,36,27,35)

#combine into table
data_68 <- data.frame(A_68, B_68, I_68, II_68, III_68, IV_68, V_68, VI_68)
data_68

#build ANOVA table

y_68 <- c(I_68, II_68, III_68, IV_68, V_68, VI_68)
 
factorA_68 <- as.factor(rep(A_68,6))
factorB_68 <- as.factor(rep(B_68,6))
g_68 <- lm(y_68 ~ factorA_68 + factorB_68 + factorA_68*factorB_68)
anova(g_68)

#looks like A does impact the response and B does not. the interaction effect is also significant

#compute sums for each combination
sums_68 <- apply(data_68[,3:8], 1, sum)
names(sums_68) <- c("(1)", "(a)", "(b)", "(ab)")
ybar_68 <- sums_68/6

#interaction plots
par(mfrow=c(1,2))
interaction.plot(A_68, B_68, ybar_68) #lines crossing
interaction.plot(B_68, A_68, ybar_68) #lines not crossing but not parallel
#interaction effect is indeed occurring
#time definitely matters, but you can also say that bacterial growth depends on the interaction of time with the culture

#effect estimate
A_hat_68<-(sums_68[4]+sums_68[2]-sums_68[3]-sums_68[1])/12
#the main effect of A is 9.92. This is the change in response produced by the change in the level of 'time' averaged over the levels of 'culture'
B_hat_68 <- (sums_68[4]+sums_68[3]-sums_68[2]-sums_68[1])/12
#the main effect of B is weak: -1.25. This is the change in response produced by the change in the level of 'culture' averaged over the levels of 'time'.
AB_hat_68 <- (sums_68[4]+sums_68[1]-sums_68[2]-sums_68[3])/12
#the interaction effect of AB is -3.92. This is the average difference between the effect of culture at the high level of time and the effect of culture at the low level of time. 
```

Based on the ANOVA, we can conclude that the main effect of time is statistically significant, but the effect of culture is not. There IS interaction between time and culture. This is backed up by the factor effects that we calculated, where abs(B_hat) was relatively low compared to the absolute values of the other effects. Now I will analyze the residuals.

```{r}
# residual analysis
par(mfrow=c(1,2))
# normal plot
qqnorm(g_68$residuals)
qqline(g_68$residuals) #this looks pretty normal
gaov_68<-aov(y_68~factorA_68+factorB_68)
shapiro.test(g_68$residuals) #it is normal, hurrah
bartlett.test(gaov_68$residuals~factorA_68) #equal variance!
bartlett.test(gaov_68$residuals~factorB_68) #equal variance!

#residual plot
plot(g_68$fitted.values,g_68$residuals) #looks homoscedastic - good
abline(h=0)
```

The assumptions are met, so our conclusions are valid. Awesome!

## Question 6.15

#### (a)

```{r}
#input data
A_615 <- as.factor(rep(c(-1,1), times=16))
B_615 <- as.factor(rep(c(-1,1), each = 2, times=8))
C_615 <- as.factor(rep(c(-1,1), each = 4, times=4))
D_615 <- as.factor(rep(c(-1,1), each = 16))
rep_615 <- as.factor(rep(1:2, each=16))
y_615 <- c(7.037,14.707,11.635,17.273,10.403,4.368,9.360,13.440,8.561,16.867,13.876,19.824,11.846,6.125,11.190,15.653, 6.376, 15.219,12.089,17.815,10.151,4.098,9.253,12.923,8.951,17.052,13.658,19.639,12.337,5.904,10.935,15.053)

data_615 <- data.frame(A_615, B_615, C_615, D_615, rep_615, y_615)

require(stats)
data_615 <- aggregate(y_615 ~ A_615 + B_615 + C_615 + D_615, FUN= mean,data=data_615)

#create linear model
g_615 <- lm(y_615~(A_615+B_615+C_615+D_615)^4,data=data_615)


#visualize effects
g_615$effects
effects_615 <- abs(g_615$effects[-1])

#plot effects
qq_615 <- qqnorm(effects_615, type="n")  # "n" means no plotting
text(qq_615$x, qq_615$y, labels = names(effects_615)) 

print(effects_615) #these are the estimated effects. looks like D is unimportant
```

#### (b) 

```{r}
anova(g_615)
```
