---
title: "Jordan_Hw6"
author: "Rachel Jordan"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 8.2

```{r}
#input data
A_82 <- rep(c(-1,1), times=8)
B_82 <- rep(c(-1,1), each = 2, times=4)
C_82 <- rep(c(-1,1), each = 4, times=2)
D_82 <- A_82*B_82*C_82 #fractional factorial, I=ABCD

A_82 <- as.factor(A_82)
B_82 <- as.factor(B_82)
C_82 <- as.factor(C_82)
D_82 <- as.factor(D_82)

y_82 <- c(7.037,14.707,11.635,17.273,10.403,4.368,9.360,13.440,8.561,16.867,13.876,19.824,11.846,6.125,11.190,15.653)

data_82 <- data.frame(A_82,B_82,C_82,D_82,y_82)

#fit the full model
g_82 <- lm(y_82 ~ (A_82+B_82+C_82+D_82)^4)
anova(g_82) #only aliases

#half-normal qq plot
effects_82 <- abs(g_82$effects[-1])
qq_82 <- qqnorm(effects_82,type="n")
text(qq_82$x, qq_82$y, labels=names(effects_82))

#fit reduced model
h_82 <- lm(y_82 ~ A_82 + B_82 + C_82 + D_82 + A_82*B_82 + A_82*C_82)
anova(h_82)

#design and interaction plots
par(mfrow=c(2,2))
plot.design(data_82)
interaction.plot(A_82,B_82,y_82)
interaction.plot(A_82,C_82,y_82)

#check assumptions
par(mfrow=c(1,2))
# normal plot
qqnorm(h_82$residuals)
qqline(h_82$residuals) #ehhhh....
haov_82<-aov(h_82)
shapiro.test(h_82$residuals) #NOT normal!!!
bartlett.test(haov_82$residuals~A_82) #homoscedastic
bartlett.test(haov_82$residuals~B_82) #homoscedastic
bartlett.test(haov_82$residuals~C_82) #homoscedastic
bartlett.test(haov_82$residuals~D_82) #homoscedastic

# residual plot
plot(h_82$fitted.values,h_82$residuals) #this looks fine
abline(h=0)
```

A:D is not significant at alpha = 0.05, but everything else is. In the reduced model, everything is significant. To get the highest mean, we want A, B, and D to be high and C to be low.

## Question 8.15

```{r}
#enter data
A_815 <- rep(c(-1,1), times=16)
B_815 <- rep(c(-1,1), each = 2, times=8)
C_815 <- rep(c(-1,1), each = 4, times=4)
D_815 <- rep(c(-1,1), each = 8, times=2)
E_815 <- A_815*B_815*C_815*D_815 #fractional factorial, I=ABCDE
y_815 <- c(7,9,34,55,16,20,40,60,8,10,32,50,18,21,44,61,8,12,35,52,15,22,45,65,6,10,30,53,15,20,41,63)

A_815 <- as.factor(A_815)
B_815 <- as.factor(B_815)
C_815 <- as.factor(C_815)
D_815 <- as.factor(D_815)

data_815 <- data.frame(A_815, B_815, C_815, D_815, E_815, y_815)

#fit the full model
g_815 <- lm(y_815 ~ (A_815+B_815+C_815+D_815+E_815)^5)
anova(g_815) #only aliases

#half-normal qq plot
effects_815 <- abs(g_815$effects[-1])
qq_815 <- qqnorm(effects_815,type="n")
text(qq_815$x, qq_815$y, labels=names(effects_815))

#fit reduced model
h_815 <- lm(y_815 ~ A_815 + B_815 + C_815 + A_815*B_815)
anova(h_815) 

#design and interaction plots
par(mfrow=c(2,2))
plot.design(data_815)
interaction.plot(A_815,B_815,y_815)
interaction.plot(B_815,A_815,y_815)

#check assumptions
par(mfrow=c(1,2))
# normal plot
qqnorm(h_815$residuals)
qqline(h_815$residuals) #this looks pretty normal
haov_815<-aov(h_815)
shapiro.test(h_815$residuals) #normal
bartlett.test(haov_815$residuals~A_815) #homoscedastic
bartlett.test(haov_815$residuals~B_815) #HETEROSCEDASTIC!!!
bartlett.test(haov_815$residuals~C_815) #homoscedastic
bartlett.test(haov_815$residuals~D_815) #homoscedastic

# residual plot
plot(h_815$fitted.values,h_815$residuals) #this looks fine
abline(h=0)
```

In the full model, only A, B, C, and A:B are significant at alpha = 0.05. All were significant in the reduced model. To increase the mean of the response, we want A and B to both be high and C to be low.

## Question 13.1

The model is y ijk = mu + tau i + beta j + (tau beta)ij + epsilon ijk where tau is part, beta is operator, tau beta is interaction between part and operator, and epsilon is random error.

#### (a)

```{r}
#input data
part <- rep(c(1,2,3,4,5,6,7,8,9,10),times=6)
operator <- rep(c(1,2),each=30)
replicate <- rep(c(1,2,3), each=10,times=2)
measurement <- c(50,52,53,49,48,52,51,52,50,47,49,52,50,51,49,50,51,50,51,46,50,51,50,50,48,50,51,49,50,49,50,51,54,48,48,52,51,53,51,46,48,51,52,50,49,50,50,48,48,47,51,51,51,51,48,50,50,50,49,48)

data_131 <- data.frame(part,operator,replicate,measurement)

#ANOVA
g_131 <- lm(measurement ~ as.factor(part) + as.factor(operator) + as.factor(part)*as.factor(operator))
anova(g_131)

#reduce model
h_131 <- lm(measurement ~ as.factor(part) + as.factor(operator))
anova(h_131)

#check assumptions
par(mfrow=c(1,2))
# normal plot
qqnorm(h_131$residuals)
qqline(h_131$residuals) #this looks pretty normal
haov_131<-aov(h_131)
shapiro.test(h_131$residuals) #normal
bartlett.test(haov_131$residuals~part) #HETEROSCEDASTIC
bartlett.test(haov_131$residuals~operator) #homoscedastic
```

There are significant differences in the variation between parts, regardless of which operator is measuring the parts. The interaction between part and operator is not significant so it was dropped to fit a reduced model. It seems that this measuring system is not very good (we are assuming here that the parts are the same size in reality).

#### (b)

```{r}
#part full model
(11.0019 - 0.6019)/6

#operator full model
(0.4167 - 0.6019)/30

#interaction full model
(0.6019 - 1.5)/3

#part reduced model
(11.0019 - 1.3350)/6

#operator reduced model
(0.4167 - 1.3350)/30

#gauge reduced model
1.3350 + 0
```

The variance component for the parts in the full model is 1.73. The variance component for the operators in the full model is -0.006 --\> 0. The variance component for the interaction term in the full model is -0.299 --\> 0. The variance component for the parts in the reduced model is 1.61. The variance component for the operators in the reduced model is -0.03 --\> 0. The variability of the gauge is smaller than the variability of the product, but not by that much. I think this is undesirable; we want the variability of the gauge to be small relative to the variability in the product, as this would indicate that the gauge is capable of distinguishing among different grades of product. That doesn't seem to be the case here.

## Question 13.5

The model is y ijk = mu + tau i + beta j + (tau beta) ij + epsilon ijk where tau is position, beta is temperature, tau beta is interaction between position and temperature, and epsilon is random error.

```{r}
#analyze problem 5.13 as a mixed model experiment

#input data
y_135 <- c(570,565,583,528,547,521,1063,1080,1043,988,1026,1004,565,510,590,526,538,532)
position <- rep(1:2,each=3,times=3)
temp <- rep(1:3, each=6)

data_135 <- data.frame(y_135,position,temp)

g_135 <- lm(y_135~(as.factor(position) + as.factor(temp))^2)
anova(g_135)

MS_position <- 7160
MS_temp <- 472671
MS_pos_temp <- 409
MSE <- 448
df_position <- 1
df_temp <- 2
df_pos_temp <- 2
dfE <- 12
#a <- 3
#b <- 20
#n <- 2

#Factor A position fixed effect
F0.position <- MS_position/MS_pos_temp
p.position <- 1-pf(F0.position,df_position,df_pos_temp)
cat("F-value for A is", F0.position,"\n")
cat("p-value for A is", p.position,"\n","\n")

# Fit the reduced model

h_135 <- lm(y_135 ~ as.factor(position) + as.factor(temp))
anova(h_135)

#check assumptions
par(mfrow=c(1,2))
# normal plot
qqnorm(h_135$residuals)
qqline(h_135$residuals) #this looks pretty normal
haov_135<-aov(h_135)
shapiro.test(h_135$residuals) #normal but barely!
bartlett.test(haov_135$residuals~position) #homoscedastic
bartlett.test(haov_135$residuals~temp) #homoscedastic
```

Position and temperature are both significant, but the interaction term is not. In the reduced model part and temperature are still significant. The assumptions are met.
