---
title: "Jordan_Hw5"
author: "Rachel Jordan"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 6.8

```{r}
#input data
A_68 <- c(-1, 1, -1,1)
B_68 <- c(-1, -1, 1, 1)
I_68 <- c(21,37,25,31)
II_68 <- c(23,38,24,29)
III_68 <- c(20,35,29,30)
IV_68 <- c(22,39,26,34)
V_68 <- c(28,38,25,33)
VI_68 <- c(26,36,27,35)

#combine into table
data_68 <- data.frame(A_68, B_68, I_68, II_68, III_68, IV_68, V_68, VI_68)
data_68

#build ANOVA table

y_68 <- c(I_68, II_68, III_68, IV_68, V_68, VI_68)
 
factorA_68 <- as.factor(rep(A_68,6))
factorB_68 <- as.factor(rep(B_68,6))
g_68 <- lm(y_68 ~ factorA_68 + factorB_68 + factorA_68*factorB_68)
anova(g_68)

#looks like A does impact the response and B does not. the interaction effect is also significant

#compute sums for each combination
sums_68 <- apply(data_68[,3:8], 1, sum)
names(sums_68) <- c("(1)", "(a)", "(b)", "(ab)")
ybar_68 <- sums_68/6

#interaction plots
par(mfrow=c(1,2))
interaction.plot(A_68, B_68, ybar_68) #lines crossing
interaction.plot(B_68, A_68, ybar_68) #lines not crossing but not parallel
#interaction effect is indeed occurring
#time definitely matters, but you can also say that bacterial growth depends on the interaction of time with the culture

#effect estimate
A_hat_68<-(sums_68[4]+sums_68[2]-sums_68[3]-sums_68[1])/12
#the main effect of A is 9.92. This is the change in response produced by the change in the level of 'time' averaged over the levels of 'culture'
B_hat_68 <- (sums_68[4]+sums_68[3]-sums_68[2]-sums_68[1])/12
#the main effect of B is weak: -1.25. This is the change in response produced by the change in the level of 'culture' averaged over the levels of 'time'.
AB_hat_68 <- (sums_68[4]+sums_68[1]-sums_68[2]-sums_68[3])/12
#the interaction effect of AB is -3.92. This is the average difference between the effect of culture at the high level of time and the effect of culture at the low level of time. 
```

Based on the ANOVA, we can conclude that the main effect of time is statistically significant, but the effect of culture is not. There IS interaction between time and culture. This is backed up by the factor effects that we calculated, where abs(B_hat) was relatively low compared to the absolute values of the other effects. Now I will analyze the residuals.

```{r}
# residual analysis
par(mfrow=c(1,2))
# normal plot
qqnorm(g_68$residuals)
qqline(g_68$residuals) #this looks pretty normal
gaov_68<-aov(y_68~factorA_68+factorB_68)
shapiro.test(g_68$residuals) #it is normal, hurrah
bartlett.test(gaov_68$residuals~factorA_68) #equal variance!
bartlett.test(gaov_68$residuals~factorB_68) #equal variance!

#residual plot
plot(g_68$fitted.values,g_68$residuals) #looks homoscedastic - good
abline(h=0)
```

The assumptions are met, so our conclusions are valid. Awesome!

## Question 6.15

#### (a)

```{r}
#input data
A_615 <- as.factor(rep(c(-1,1), times=16))
B_615 <- as.factor(rep(c(-1,1), each = 2, times=8))
C_615 <- as.factor(rep(c(-1,1), each = 4, times=4))
D_615 <- as.factor(rep(c(-1,1), each = 8, times=2))
rep_615 <- as.factor(rep(1:2, each=16))
y_615 <- c(7.037,14.707,11.635,17.273,10.403,4.368,9.360,13.440,8.561,16.867,13.876,19.824,11.846,6.125,11.190,15.653, 6.376, 15.219,12.089,17.815,10.151,4.098,9.253,12.923,8.951,17.052,13.658,19.639,12.337,5.904,10.935,15.053)

data_615 <- data.frame(A_615, B_615, C_615, D_615, rep_615, y_615)

#create linear model
g_615 <- lm(y_615~(A_615+B_615+C_615+D_615)^4,data=data_615)


#visualize effects
g_615$effects
effects_615 <- abs(g_615$effects[-1])

print(effects_615) #these are the estimated effects. looks like interactions with D are unimportant
```

#### (b)

```{r}
anova(g_615)
```

Everything except the interaction terms that contain D (grain refiner) affect cracking (they are significant at alpha = 0.05).

#### (c)

The regression model is y-hat = 11.99 + (8.54/2)x1 + (11.25/2)x2 + (10.17/2)x3 + (5.54/2)x4 + (5.47/2)x1x2 + (11.34/2)x1x3 + (8.87/2)x1x2x3

```{r}
mean(y_615)

#reduced model
g_reduced_615 <- lm(y_615~A_615+B_615+C_615+D_615 + A_615*B_615 + A_615*C_615 + A_615*B_615*C_615)
anova(g_reduced_615)
abs(g_reduced_615$effects)
```

#### (d) 

```{r}
#residual analysis


par(mfrow=c(1,2))
# normal plot
qqnorm(g_reduced_615$residuals)
qqline(g_reduced_615$residuals) #normal
gaov_615<-aov(y_615~A_615+B_615+C_615+D_615 + A_615*B_615 + A_615*C_615 + A_615*B_615*C_615)
shapiro.test(g_reduced_615$residuals) #normal!
bartlett.test(gaov_615$residuals~A_615) #homoscedastic for A
bartlett.test(gaov_615$residuals~B_615) #omoscedastisitic for B
bartlett.test(gaov_615$residuals~C_615) #homoscedastic
bartlett.test(gaov_615$residuals~D_615) #homoscedastic

# residual plot
plot(g_615$fitted.values,g_615$residuals) #this looks fine
abline(h=0)
```

#### (e) 

```{r}
bartlett.test(gaov_615$residuals~A_615) 
bartlett.test(gaov_615$residuals~B_615)
bartlett.test(gaov_615$residuals~C_615) 
bartlett.test(gaov_615$residuals~D_615) 
```

There is no indication that the variables affect the variability because the variance is not significantly different with any of the significant factors.

#### (f) 

```{r}
par(mfrow=c(2,2))
plot.design(data_615) 
```

To result in smaller cracks, we want A, B, and D to be low and C to be high.

## Question 6.26

```{r}
#enter data
A_626 <- as.factor(rep(c(-1,1), times=16))
B_626 <- as.factor(rep(c(-1,1), each = 2, times=8))
C_626 <- as.factor(rep(c(-1,1), each = 4, times=4))
D_626 <- as.factor(rep(c(-1,1), each = 8, times=2))
E_626 <- as.factor(rep(c(-1,1), each = 16))
y_626 <- c(7,9,34,55,16,20,40,60,8,10,32,50,18,21,44,61,8,12,35,52,15,22,45,65,6,10,30,53,15,20,41,63)

data_626 <- data.frame(A_626, B_626, C_626, D_626, E_626, y_626)
```

#### (a) 

```{r}
#create linear model
g_626 <- lm(y_626~(A_626+B_626+C_626+D_626+E_626)^5,data=data_626)

#visualize effects
g_626$effects
effects_626 <- abs(g_626$effects[-1])

qq_626 <- qqnorm(effects_626, type="n")  # "n" means no plotting
text(qq_626$x, qq_626$y, labels = names(effects_626))
```

The effect of B appears to be large, and A, C, and A\*B also seem important.

#### (b) 

```{r}
anova(g_626) #not enough df's to run this fully so I'm just going to toss out some terms based on the half-normal probability plot.

g_reduced_626 <- lm(y_626 ~ A_626 + B_626 + C_626 + A_626*B_626) #these are all significant! yippee
anova(g_reduced_626)
```

#### (c) 

```{r}
mean(y_626)

abs(g_reduced_626$effects[-1])
```

The model is y_hat = 30.53 + (33.4/2)x1 + (95.99/2)x2 + (27.4/2)x3 + (22.45/2)x1x2.

#### (d) 

```{r}
#residual analysis

par(mfrow=c(1,2))
# normal plot
qqnorm(g_reduced_626$residuals)
qqline(g_reduced_626$residuals) #normal
gaov_626<-aov(y_626 ~ A_626 + B_626 + C_626 + A_626*B_626)
shapiro.test(g_reduced_626$residuals) #normal!
bartlett.test(gaov_626$residuals~A_626) #homoscedastic for A
bartlett.test(gaov_626$residuals~B_626) #uh-oh! heterscedastic for B!
bartlett.test(gaov_626$residuals~C_626) #homoscedastic for C

```

#### (e) 

```{r}
# residual plot
plot(g_626$fitted.values,g_626$residuals) #this looks fine
abline(h=0)

plot(g_626$fitted.values,g_626$A_626) 
plot(g_626$fitted.values,g_626$B_626) 
plot(g_626$fitted.values,g_626$C_626) 
plot(g_626$fitted.values,g_626$D_626) 
plot(g_626$fitted.values,g_626$E_626) 
```

The points are all on the line in the residuals vs fitted plot. They look kind of randomly distributed in the plots of the factors against the residuals. Seems good.

#### (f) 

```{r}
#interaction plots
par(mfrow=c(1,2))
interaction.plot(A_626, B_626, y_626) #lines not crossing but not parallel
interaction.plot(B_626, A_626, y_626) #lines not crossing but not parallel
#interaction effect is occurring between A and B. The interaction of aperture and exposure time affects the yield.
```

A\*B is significant and this interaction affects the yield.

#### (g)

```{r}
par(mfrow=c(2,2))
plot.design(data_626) 
```

We want A, B, and C to be high to produce a maximum yield. D and E are insignificant.

## Question 7.3 

```{r}
#input data
A_73 <- as.factor(rep(c(-1,1), times=8))
B_73 <- as.factor(rep(c(-1,1), each = 2, times=4))
C_73 <- as.factor(rep(c(-1,1), each = 4, times=2))
D_73 <- as.factor(rep(c(-1,1), each = 8))
Block1_73 <- c(7.037,14.707,11.635,17.273,10.403,4.368,9.360,13.440,8.561,16.867,13.876,19.824,11.846,6.125,11.190,15.653)
Block2_73 <- c(6.376, 15.219,12.089,17.815,10.151,4.098,9.253,12.923,8.951,17.052,13.658,19.639,12.337,5.904,10.935,15.053)

data_73 <- data.frame(A_73, B_73, C_73, D_73, Block1_73, Block2_73)

y_73 <- c(Block1_73,Block2_73)

```

## Question 7.7

```{r}
#ABCDE confounded with 2 blocks
#enter data
A_77 <- rep(c(-1,1), times=16)
B_77 <- rep(c(-1,1), each = 2, times=8)
C_77 <- rep(c(-1,1), each = 4, times=4)
D_77 <- rep(c(-1,1), each = 8, times=2)
E_77 <- rep(c(-1,1), each = 16)
y_77 <- c(7,9,34,55,16,20,40,60,8,10,32,50,18,21,44,61,8,12,35,52,15,22,45,65,6,10,30,53,15,20,41,63)

ABCDE_77 <- A_77*B_77*C_77*D_77*E_77 #want to confound ABCDE
data_77 <- data.frame(factor(A_77), factor(B_77), factor(C_77), factor(D_77), (E_77), ABCDE_77, y_626)
#ABCDE = blocks

g_77 <- lm(y_77 ~ (A_77+B_77+C_77+D_77)^4)
anova(g_77) #looks like A, B, C, and AB are significant

Blocks_77 <- ABCDE_77
h_77 <- lm(y_626 ~ A_626 + B_626 + C_626 + A_626*B_626 + Blocks_77)
anova(h_77) #A, B, C, and AB are significant

#residual analysis

par(mfrow=c(1,2))
# normal plot
qqnorm(h_77$residuals)
qqline(h_77$residuals) #normal
gaov_77<-aov(y_77 ~ A_77 + B_77 + C_77 + A_77*B_77)
shapiro.test(h_77$residuals) #normal!
bartlett.test(gaov_77$residuals~A_77) #homoscedastic for A
bartlett.test(gaov_77$residuals~B_77) #uh-oh! heterscedastic for B!
bartlett.test(gaov_77$residuals~C_77) #homoscedastic for C
bartlett.test(gaov_77$residuals~Blocks_77) #homoscedastic
```

## Question 7.9 

```{r}
#4 blocks with ACDE and BCD (and consequently ABE) confounded


```
